---
title: "Modeling with tidymodels in R - DataCamp"
author: "Edneide Ramalho"
date: "`r format(Sys.time(), '%d/%m/%Y')`"
output:
  html_document:
    highlight: textmate
    logo: logo.png
    theme: jou
    number_sections: yes
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: no
    df_print: paged
  word_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, echo=FALSE}
htmltools::img(src = knitr::image_uri("img/tidymodels.png"), 
               alt = 'logo', 
               style = 'position:absolute; top:0; right:0; padding:10px;',
               width = "200",
               heigth = "200")
```

# Tracks for this course

-   Machine Learning Fundamentals in R

-   Machine Learning Scientist with R

-   Supervised Machine Learning in R

# Imports

```{r packages}
library(tidyverse)
library(tidymodels)
```

# The tidymodels ecosystem

## Collection of machine learning packages

-   Data resampling: `rsample`

-   Feature engineering: `recipes`

-   Model fitting: `parsnip`

-   Model Tuning: `tune` e `dials`

-   Model evaluation: `yardstick`

## Data resampling 

Create training and tests sets

-   Guards against **overfitting**

-   Common ratio

    -   75% training

    -   25% test

**Training data**

-   Feature engineering

-   Model fitting and tunning

**Test data**

-   Estimate model performance on new data

## Fuel efficiency data

-   Dataset: `mpg`

-   Outcome variable: `hwy` (highway fuel efficiency in miles per gallon - mpg)

```{r data}
glimpse(mpg)
```

## Data resampling with tidymodels
