---
title: "Modeling with tidymodels in R - DataCamp"
author: "Edneide Ramalho"
date: "`r format(Sys.time(), '%d/%m/%Y')`"
output:
  html_document:
    highlight: textmate
    logo: logo.png
    theme: jou
    number_sections: yes
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: no
    df_print: paged
  word_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r, echo=FALSE}
htmltools::img(src = knitr::image_uri("img/tidymodels.png"), 
               alt = 'logo', 
               style = 'position:absolute; top:0; right:0; padding:10px;',
               width = "200",
               heigth = "200")
```

# Imports

```{r packages}
library(tidyverse)
library(tidymodels)
```

# Classification models

```{r}
# data
leads_df <- readRDS('datasets/leads_df.rds')
glimpse(leads_df)
```

## Data resampling

```{r}
leads_split <- initial_split(leads_df,
                             prop = 0.75,
                             strata = purchased)

leads_training <- leads_split %>% 
  training()

leads_test <- leads_split %>%
  testing()

```

## Logistic regression model specification

```{r}
logistic_model <- logistic_reg() %>% 
  set_engine('glm') %>% 
  set_mode('classification')
```

## Model fitting

```{r}
logistic_fit <- logistic_model %>% 
  fit(purchased ~ total_visits + total_time,
      data = leads_training)
```

## Predicting outcome categories

```{r}
class_preds <- logistic_fit %>% 
  predict(new_data = leads_test, 
          type = 'class')

class_preds
```

## Estimated probabilities 

```{r}
prob_preds <- logistic_fit %>% 
  predict(new_data = leads_test,
          type = 'prob')

prob_preds
```

## Combining results

```{r}
leads_results <- leads_test %>% 
  select(purchased) %>% 
  bind_cols(class_preds, prob_preds)

leads_results
```

# Exercises - part 1

> 1.  **Data resampling**

The first step in a machine learning project is to create training and test datasets for model fitting and evaluation. The test dataset provides an estimate of how your model will perform on new data and helps to guard against overfitting.

You will be working with the `telecom_df` dataset which contains information on customers of a telecommunications company. The outcome variable is `canceled_service` and it records whether a customer canceled their contract with the company. The predictor variables contain information about customers' cell phone and Internet usage as well as their contract type and monthly charges.

The `telecom_df` tibble has been loaded into your session.

-   Create an `rsample` object, `telecom_split`, that contains the instructions for randomly splitting the `telecom_df` data into training and test datasets.

    -   Allocate 75% of the data into training and stratify the results by `canceled_service`.

-   Pass the `telecom_split` object to the appropriate `rsample` functions to create the training and test datasets.

-   Check the number of rows in each datasets by passing them to the `nrow()` function.

```{r}
# data
telecom_df <- readRDS('datasets/telecom_df.rds')
glimpse(telecom_df)

# Create data split object
telecom_split <- initial_split(telecom_df, prop = 0.75,
                     strata = canceled_service)

# Create the training data
telecom_training <- telecom_split %>% 
  training()

# Create the test data
telecom_test <- telecom_split %>% 
  testing()

# Check the number of rows
paste0("#Rows training: ", nrow(telecom_training))
paste0("#Rows test: ", nrow(telecom_test))
```

# Fitting a logistic regression model

In addition to regression models, the `parsnip` package also provides a general interface to classification models in R.

In this exercise, you will define a `parsnip` logistic regression object and train your model to predict `canceled_service` using `avg_call_mins`, `avg_intl_mins`, and `monthly_charges` as predictor variables from the `telecom_df` data.

-   Initialize a logistic regression object, `logistic_model`, with the appropriate `parsnip` function.

-   Use the `'glm'` engine.

-   Set the mode to `'classification'`.

-   Print the `logistic_model` object to view its specification details.

-   Train your model to predict `canceled_service` using `avg_call_mins`, `avg_intl_mins`, and `monthly_charges` as predictor variables from the `telecom_training` dataset.

```{r}
# Specify a logistic regression model
logistic_model <- logistic_reg() %>% 
  set_engine('glm') %>% 
  set_mode('classification')

# Print the model specification
logistic_model
```

```{r}
# Fit to training data
logistic_fit <- logistic_model %>% 
  fit(canceled_service ~ avg_call_mins + avg_intl_mins + monthly_charges, data = telecom_training)

# Print model fit object
logistic_fit
```

> 2.  **Combining test dataset results**

Evaluating your model's performance on the test dataset gives insights into how well your model predicts on new data sources. These insights will help you communicate your model's value in solving problems or improving decision making.

Before you can calculate classification metrics such as sensitivity or specificity, you must create a results tibble with the required columns for `yardstick` metric functions.

In this exercise, you will use your trained model to predict the outcome variable in the `telecom_test` dataset and combine it with the true outcome values in the `canceled_service` column.

-   Use your trained model and the `predict()` function to create a tibble, `class_preds`, with predicted outcome variable categories using the test dataset.

```{r}
# Predict outcome categories
class_preds <- predict(logistic_fit, new_data = telecom_test,
                       type = 'class') 
class_preds
```

-   Now create a tibble, `prob_preds`, with the estimated probabilities for each category in the outcome variable using the test dataset.

```{r}
# Obtain estimated probabilities for each outcome value
prob_preds <- predict(logistic_fit, new_data = telecom_test,
                       type = 'prob') 
prob_preds
```

-   Select the outcome variable from the `telecom_test` data.

-   Add the `class_preds` and `prob_preds` tibbles along the column axis.

```{r}
telecom_results <- telecom_test %>% 
  select(canceled_service) %>% 
  bind_cols(class_preds, prob_preds)
```

# Assessing model fit 

In `tidymodels` outcome variables needs to be a factor

-   First level is positive class

-   Check order with `levels()`

```{r}
leads_df %>% glimpse()

levels(leads_df$purchased)
```

## Confusion matrix

```{r, echo=FALSE}
# htmltools::img(src = knitr::image_uri("img/confusion_matrix.png"),
#                width = "200",
#                heigth = "200",
#                align = 'center')
```

```{r, echo=FALSE}
library(htmltools)
# Replace "your_image.png" with the actual filename of your image
img_path <- "img/confusion_matrix.png"

# Create the image container with CSS styles to center it
img_centered <- div(
  style = "display: flex; justify-content: center;",
  img(src = img_path, width = "350")
)
img_centered
```

**Correct predictions**

-   True Positive (TP)

-   True Negative (TN)

**Classifcation Errors**

-   False Positive (FP)

-   False Negative (FN)

```{r}
conf_mat(leads_results,
         truth = purchased,
         estimate = .pred_class)
```

## Classification accuracy

$$
\dfrac{TP + TN}{TP + TN + FP + FN}
$$

```{r}
accuracy(leads_results, 
         truth = purchased,
         estimate = .pred_class)
```

## Sensitivity

In many cases accuracy is not the best metric.

-   `leads_df` data

    -   Classifying all as 'no' gives 64% accuracy

**Sensitivity**

Proportion of all positive cases that were correctly classified.

-   Of custumers who did purchased, what proportion did our model predicted correctly?

    -   Lower false negatives increase sensitivity.

$$
\dfrac{TP}{TP + FN}
$$

```{r}
sens(leads_results, 
         truth = purchased,
         estimate = .pred_class)
```

## Specificity

$$
\dfrac{TN}{TN + FP}
$$

Is the proportion of all negative cases that were correctly classified.

-   Of customers who did not purchased, what proportion did our model predicted correctly?

    -    Lower false positive increases specificity

$1 - \text{specificity}$ : False Positive Rate (FPR)

-   Proportion of false positives among True negatives

```{r}
spec(leads_results, 
         truth = purchased,
         estimate = .pred_class)
```

## Creating a metric set

```{r}
custom_metrics <- metric_set(accuracy, sens, spec)

custom_metrics(leads_results,
               truth = purchased,
               estimate = .pred_class)
```

## Many metrics

-   `accuracy()`, `kap()`, `sens()`, `spec()`, `ppv()`, `npv(),` `mcc()`, `j_index()`, `bal_accuracy()`, `detection_prevalence()`, `precision()`, `recall()`, `f_means()`

-   Pass the results of conf_mat() to summary() to calculate all

-   Documentation: <https://yardstick.tidymodels.org/reference>

# Exercises - part 1

> 1.  **Evaluating performance with yardstick**

In the previous exercise, you calculated classification metrics from a sample confusion matrix. The `yardstick` package was designed to automate this process.

For classification models, `yardstick` functions require a tibble of model results as the first argument. This should include the actual outcome values, predicted outcome values, and estimated probabilities for each value of the outcome variable.

In this exercise, you will use the results from your logistic regression model, `telecom_results`, to calculate performance metrics.

-   Use the appropriate `yardstick` function to create a confusion matrix using the `telecom_results` tibble.

    ```{r}
    # Calculate the confusion matrix
    conf_mat(telecom_results, truth = canceled_service,
        estimate = .pred_class)
    ```

-   Calculate the **accuracy** of your model with the appropriate `yardstick` function.

    ```{r}
    # Calculate the accuracy
    accuracy(telecom_results, truth = canceled_service,
        estimate = .pred_class)
    ```

-   Calculate the **sensitivity** of your model.

    ```{r}
    # Calculate the sensitivity
    sens(telecom_results, truth = canceled_service,
        estimate = .pred_class)
    ```

-   Calculate the **specificity** of your model.

    ```{r}
    # Calculate the specificity
    spec(telecom_results,
                truth = canceled_service, 
                estimate = .pred_class)
    ```

> **2. Creating custom metric sets**

The `yardstick` package also provides the ability to create custom sets of model metrics. In cases where the cost of obtaining false negative errors is different from the cost of false positive errors, it may be important to examine a specific set of performance metrics.

Instead of calculating accuracy, sensitivity, and specificity separately, you can create your own metric function that calculates all three at the same time.

In this exercise, you will use the results from your logistic regression model, `telecom_results`, to calculate a custom set of performance metrics. You will also use a confusion matrix to calculate all available binary classification metrics in `tidymodels`all at once.

```{r}
# Create a custom metric function
telecom_metrics <- metric_set(accuracy, sens, spec)

# Calculate metrics using model results tibble
telecom_metrics(telecom_results, truth = canceled_service,
                estimate = .pred_class)

# Create a confusion matrix
conf_mat(telecom_results,
         truth = canceled_service,
         estimate = .pred_class) %>% 
  # Pass to the summary() function
  summary()
```
